Patient-Specific Neural Field Model

### 2010-07-19 ltx/data_paper/DataPaper.tex
continued to add model derivation up to Euler method

\subsection{Neural Field Model}
The neural field equations used as the parametric form of the model are based on the influential papers from Wilson and Cowan~\cite{Wilson1973}, and Amari~\cite{Amari1977}. This class of model describes a continuous cortical sheet or surface, relating mean firing rates of pre and post-synaptic neural populations. The neural field, $v\left( {\mathbf{r},t} \right)$, at position $\mathbf{r}$ and time $t$, is the aggregated post-synaptic potentials described by the temporal convolution 
\begin{equation}
	\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}, 
\end{equation}   
where $h(\cdot)$ is the synaptic response kernel and $g(\cdot)$ describes the mean firing rate. The firing rate is described by spatial convolution
\begin{equation}
	\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}, 
\end{equation}
where $f(\cdot)$ relates mean post-synaptic potentials to mean firing rates, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing the cortical sheet or surface. The firing rate function is the sigmoidal activation function described by
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{1}{1 + \exp \left( \varsigma \left( v_0 - v\left(\mathbf{r}',t\right) \right) \right)}, 
\end{equation}
where $v_0$ and $\varsigma$ describe the firing threshold the slope of the sigmoid respectively. By substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
	\int_{-\infty}^t 
	h\left(t - t'\right) \int_\Omega
	w\left(\mathbf{r},\mathbf{r}'\right) 
	f\left( v\left( \mathbf{r}',t' \right)\right)
	\textrm{d}\mathbf{r}'dt'.
\end{equation}
To arrive at the general intergral-differential equation form of the model, we shall express the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}
	\label{FinalFormContinuous} 
	Dv\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)\textrm{d}\mathbf{r}'}. 
\end{equation}
\subsection{Data Collection and Pre-processing}
\begin{itemize}
	\item Show medical imaging
	\item Show LFP Data (background, seizure, post-seizure)
	\item Show spectral properties of LFPs
\end{itemize}

>>>>

\subsection{Neural Field Model}
The neural field equations used as the parametric form of the model are based on the influential papers from Wilson and Cowan~\cite{Wilson1973}, and Amari~\cite{Amari1977}. This class of model describes a continuous cortical sheet or surface, relating mean firing rates of pre and post-synaptic neural populations. The neural field, $v\left( {\mathbf{r},t} \right)$, at position $\mathbf{r}$ and time $t$, is the aggregated post-synaptic potentials described by the temporal convolution 
	\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}, 
where $h(\cdot)$ is the synaptic response kernel and $g(\cdot)$ describes the mean firing rate. Assuming an infinite propagation velocity for action potentials within the field, the incoming firing rate the position $\mathbf{r}$ is described by spatial convolution
	\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}, 
where $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing the cortical sheet or surface. The function $f(\cdot)$ relates mean post-synaptic potentials to mean firing rates and follows a sigmoid described by
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{1}{1 + \exp \left( \varsigma \left( v_0 - v\left(\mathbf{r}',t\right) \right) \right)}, 
where $v_0$ and $\varsigma$ describe the firing threshold the slope of the sigmoid respectively. By substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} we get the spatiotemporal model 
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
	\int_{-\infty}^t 
	h\left(t - t'\right) \int_\Omega
	w\left(\mathbf{r},\mathbf{r}'\right) 
	f\left( v\left( \mathbf{r}',t' \right)\right)
	\textrm{d}\mathbf{r}'dt'.
To arrive at the general integral-differential equation form of the model, we shall express the synaptic response kernel as a Green's function 
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
where $D$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
	\label{FinalFormContinuous} 
	Dv\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)\textrm{d}\mathbf{r}'}. 
By assuming the synaptic response kernel has the form 
	\label{SynapticRespKernel} h(t) = \eta(t)\exp{\left(-\zeta t\right)}, 
where $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function, the operator $D$ is
	D=\frac{d}{dt} + \zeta,
yielding
	\label{FinalFormContinuous} 
	\frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)\textrm{d}\mathbf{r}'}. 
The neural field equations must be written as a discrete-time finite-dimensional model in order to relate it to patient-specific data. The discrete time model is found applying the a first-order Euler method giving
\begin{equation}\label{EulerMethod}
	\frac{v\left( \mathbf{r},t+T_s \right)-v\left( \mathbf{r},t \right)}{T_s} + \zeta v\left(\mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right) f\left( {v\left( \mathbf{r}',t \right)}\right)\textrm{d}\mathbf{r}'},
where $T_s$ is the time step or sampling period. Rearranging equation~\ref{EulerMethod} gives the integro-difference equation
	\label{DiscreteTimeModel} 
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'}, 
where $\xi = 1-T_s\zeta$.

\subsection{Data Collection and Pre-processing}
\begin{itemize}
	\item Show medical imaging
	\item Show LFP Data (background, seizure, post-seizure)
	\item Show spectral properties of LFPs
\end{itemize}
